Triangulating iconicity: Coding analysis

### 2024-04-03 paper_02_main_analyses.Rmd
cleanup and generate github md for main analyses

# custom functions
source("TI_functions.R")
```

@@

# get consensus coding data
d = read_excel("data/ideophones_coded.xlsx") %>% arrange(filename)

@@

d.scores = read_excel("data/ideophones_guessability.xlsx") %>%
  dplyr::select(-category)
d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))

# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
d <- d %>%
  group_by(study) %>%
  mutate(logodds = probitlink(score))

# add Z score to make scores more comparable in plots across studies
d <- d %>%
  group_by(study) %>%
  mutate(score_z = scale(score,center=T,scale=T))

# get ratings data
d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") %>%
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language")) |>
  mutate(filename = gsub("_org","",filename))

@@

# also write a simplified copy for easy inspection in Excel (helpful when writing)
d %>%
  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,rating,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
  write_xlsx(path="data/ideophones_coded_guessed_rated_simplified.xlsx")

ideophones_filename <- d |> ungroup() |> select(ideophone,filename) |>
  mutate(filename = str_to_lower(filename))

# load the full ratings data for some of the reporting in the paper
d.ratings.full <- read_xlsx("data\\ideophones_rated.xlsx") |>
  mutate(list = gsub(".*([A-Z])$", "\\1",item)) |>
  mutate(filename = gsub("^(.*)_.*$", "\\1", str_to_lower(item))) |>
  mutate(filename = gsub("Coll_","", filename)) |>
  select(-item)

@@

## Figures
Here we generate some descriptive stats and create the main figures for the paper.

@@

d %>%
  group_by(study) %>%
  dplyr::summarize(max=max(score),min=min(score),mean=mean.na(score))

@@

### Figure: Overview panel
The overview panel is created separately but we use the data from the guessability studies as a canvas.

@@

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

@@

ggsave("figures\\fig2-guessability-blank.pdf",height=5,width=7.5)

Let's plot the guessability by study and iconicity rating

```{r figures_paper_guessability_rating}

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
  NULL

pB <- ggplot(data=dp, aes(x=rating,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") +
  labs(title="Guessability (both studies) by iconicity rating",
       x="iconicity rating",
       y="") +
  geom_point(size=2.0,shape=21) +
  geom_smooth(method="lm") +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
#ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\fig3-panelAB.png",height=4,width=9,bg="white")

@@

# simplest thing we can do is a correlation.

cor.test(d$logodds,d$C_cumulative,method="pearson")

d.Collabra <- d %>% filter(study == "Collabra")
d.Language <- d %>% filter(study == "Language")
cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")
cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")

p0 <- cor.test(d$logodds,d$C_cumulative,method="pearson")$p.value
p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")$p.value
p2 <- cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")$p.value
p.adjust(c(p0,p1,p2),method="bonferroni")
pvalues <-p.adjust(c(p0,p1,p2),method="bonferroni")

@@

@@

dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = forcats::fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

# mean of C_cumuluative
dp.categories |>
  group_by(category) |>
  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

@@

# binary comparison
dp.categories |>
  ungroup() |>
  group_by(category_binary) %>%
  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
  arrange(-mean_C)

@@

# get means and standard deviations across all ideophones and for each language and category
mean(d.allratings$rating)
  # 2.948339
sd(d.allratings$rating)

@@

```{r linear_regressions}

@@

A simple correlation test suggests a positive relation between iconicity rating and cumulative iconicity. As above, we can go a bit deeper by using mixed effects modelling. We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away.

@@

# simple correlation test
cor.test(d$rating_z,d$C_cumulative,method="pearson")

@@

@@

As above, we can assess the relative weight of different features using model comparison. We build models by adding the congruency measures in rank order of attestation in the coded data. Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance. Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (log likelihood difference = 21.3).

@@

@@

>>>>

@@

# get consensus coding data
d = read_excel("data/ideophones_coded.xlsx") |> arrange(filename)

@@

d.scores = read_excel("data/ideophones_guessability.xlsx") |>
  dplyr::select(-category)
d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))

# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
d <- d |>
  group_by(study) |>
  mutate(logodds = probitlink(score))

# add Z score to make scores more comparable in plots across studies
d <- d |>
  group_by(study) |>
  mutate(score_z = scale(score,center=T,scale=T))

# get ratings data
d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") |>
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language")) |>
  mutate(filename = gsub("_org","",filename))

@@

# load and merge the full ratings data for some of the reporting in the paper
ideophones_filename <- d |> ungroup() |> select(ideophone,filename) |>
  mutate(filename = str_to_lower(filename))

d.ratings.full <- read_xlsx("data/ideophones_rated.xlsx") |>
  mutate(list = gsub(".*([A-Z])$", "\\1",item)) |>
  mutate(filename = gsub("^(.*)_.*$", "\\1", str_to_lower(item))) |>
  mutate(filename = gsub("Coll_","", filename)) |>
  select(-item)

@@

## Figures
Here we generate some descriptive stats and create the main figures for the paper. Note that Figures 1 and 2 in the paper are illustrative and their final versions are not generated in R.

@@

d |>
  group_by(study) |>
  dplyr::summarize(max=max(score),min=min(score),mean=mean.na(score))

@@

### Figure 2: Overview panel
The overview of the triangulation method is created separately in CorelDraw but we use the data from the guessability studies as a canvas.

@@

# make a copy dp of the data in which we convert some measures to factors for easy plotting & faceting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

@@

ggsave("figures/fig2-guessability-blank.pdf",height=5,width=7.5)

@@

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures/fig3-panelAB.png",height=4,width=9,bg="white")

@@

# Pearson correlation for both of the studies
d.Collabra <- d |> filter(study == "Collabra")
d.Language <- d |> filter(study == "Language")
correlation.Collabra <- cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")
correlation.Language <- cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")

# Bonferroni-correcting the p values for multiple comparisons
p1 <- correlation.Collabra$p.value
p2 <- correlation.Language$p.value
pvalues <-p.adjust(c(p1,p2),method="bonferroni")

@@

The correlation is found in the Collabra study (Pearson’s r = `r round(correlation.Collabra$estimate,2)`, p = `r pvalues[1]`) as well as in the Language study (Pearson’s r = `r round(correlation.Language$estimate,2)`, p `r pvalues[2]`, p-values Bonferroni-corrected for multiple comparisons).

@@

dp.categories <- d |>
  filter(category %notin% "Other") |>
  mutate(category = forcats::fct_reorder(category, -C_cumulative, .fun='mean')) |>
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) |>
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) |> # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

# mean of C_cumuluative
dp.categories |>
  group_by(category) |>
  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
  arrange(-mean_C)

@@

# binary comparison
dp.categories |>
  ungroup() |>
  group_by(category_binary) |>
  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
  arrange(-mean_C)

@@

# get means and standard deviations across all ideophones and for each language and category
mean(d.ratings.full$rating)

sd(d.ratings.full$rating)

@@

@@

```{r coding_ratings_correlation}
correlation_coding_ratings <- cor.test(d$rating_z,d$C_cumulative,method="pearson")
```

A simple correlation test suggests a positive relation between iconicity rating and cumulative iconicity: Pearson's *r* = `r round(correlation_coding_ratings$estimate,2)`, p = `r correlation_coding_ratings$p.value`.

As above, we can go a bit deeper by using mixed effects modelling. We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away.

@@

@@

m0m1.chisquare <- round(anova(m0,m1)$Chisq[2],2)
m0m1.pvalue <- anova(m0,m1)$"Pr(>Chisq)"[2]
m0m1.loglikdiff <- abs(anova(m0,m1)$logLik[1] - anova(m0,m1)$logLik[2])

@@

We find that a model with cumulative iconicity as an additional fixed effect achieves a fit that is significantly better than a null model without it (χ2(1) = `r m0m1.chisquare`, p = `r m0m1.pvalue`, log likelihood difference `r m0m1.loglikdiff`). The fixed effect of cumulative iconicity is estimated at `r fixef(m1)[["C_cumulative"]]`.

```{r ratings_stats_3}

@@

anova(m0,m3) # aspect helps
anova(m3,m4) # magnitude doesn't additionally help
anova(m4,m5) # modality helps
anova(m3,m5) # modality helps also relative to m3
anova(m5,m6) # length helps
anova(m6,m7) # irregular doesn't additionally help
anova(m5,m8) # aspect + modality + length is where it's at
# so m8 is the best performing most complex model
anova(m0,m8)

m0m8.chisquare <- round(anova(m0,m8)$Chisq[2],2)
m0m8.pvalue <- anova(m0,m8)$"Pr(>Chisq)"[2]
m0m8.loglikdiff <- abs(anova(m0,m8)$logLik[1] - anova(m0,m8)$logLik[2])

m8.estimates <- summary(m8)[["coefficients"]][,1]

fixef(m8)

@@

We can also assess the relative weight of different features using model comparison. We build models by adding the congruency measures in rank order of attestation in the coded data. Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance. Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (Chisquare = `r m0m8.chisquare`, p = `r m0m8.pvalue`, log likelihood difference = `r m0m8.loglikdiff`).

The model estimates given in the output of `fixef(m8)` represent the contribution of a given fixed effect while controlling for others, but they are NOT additive (i.e. they cannot be summed to arrive at the effect of `C_cumulative`).

* `C_length`: `r round(m8.estimates["C_length"],2)`
* `C_modality`: `r round(m8.estimates["C_modality"],2)`
* `C_aspect`: `r round(m8.estimates["C_aspect"],2)`
